Architectures for workflow integration in science gateways

### 2016-04-11 paper.tex
Updates in section 3

\usepackage{svg}

\newcommand{\todo}[1]{\color{blue} #1 \color{black}}
\newcommand{\note}[1]{\color{blue}\textbf{Note: }\textit{#1} \color{black}}

@@

@@

\begin{figure}
\centering
\def\svgwidth{0.5\columnwidth}
\input{images/tight.pdf_tex}
\label{archi:tight}
\caption{Tight integration}
\end{figure}

@@

Or as a component.

= Examples
* VIP
* Silvia's gateway, starting from VBrowser

P-Grade: http://onlinelibrary.wiley.com/doi/10.1002/cpe.1654/full

= Description 
*  Workflow engine is deployed as a (Web) service called by the
science gateway. Probably the most adopted architecture.

= Evaluation
+ Easy to implement in the SG (just call the service)
- Load-balancing between services is difficult
(refer to VIP's experience).
- Need to wrap workflow engines in
services.
- Need to know where to submit each workflow, e.g. when
multiple workflows are involved.
- The science gateway has no
knowledge of the detailed progress of the workflow, in particular
job-level information (statuses, stdout, stderr). If we wanted to do
that, specific feedback channels have to be developed, which requires more work (see VIP).

\begin{figure}
\centering
\def\svgwidth{0.5\columnwidth}
\input{images/service.pdf_tex}
\label{archi:service}
\caption{Service integration}
\end{figure}

@@

Examples:
* CBRAIN and Niak, CBRAIN and FSL, CBRAIN and Nipype if time permits

+ Simplified parallelization
- Needs extension on the workflow engine side. Simple if workflow engine already supports task submission.

In  this model, the task of the SG is only to submit command line tools and command line tools submitted by these command line tools. The architecture is really upsided-down (was: workflow engine on top, jobs below. Is: job on top, workflow engine below). And this is what it makes it so powerful.

Sub-tasking also requires dependencies between tasks.

Can support various models such as pilot jobs, fsl sub, etc.

No global vision of the workflow.

\subsection{Agent model}

= Examples
* SHIWA Pool

@@

>>>>

\usepackage{svg}
\usepackage{url}
\usepackage{xspace}

\newcommand{\todo}[1]{\color{blue}\xspace\emph{#1}\xspace\color{black}}
\newcommand{\note}[1]{\color{blue}\textbf{Note: }\textit{#1} \color{black}}

@@

c. Data movements can be triggered by the user, to upload input data
or download processed data, or by the workflow engine, to transfer
data across the infrastructure.

@@

See Figure~\ref{archi:tight}. The workflow engine is tightly
integrated with the science gateway, which means that it shares code,
framework components, libraries, etc. For instance, the workflow
engine might be a portlet in a Liferay portal, a controller in a Ruby
on Rails application, and so on. The workflow engine and the science
gateway usually share a database where application, users and other
resources are stored. In this model, task control and data control are
initiated from the science gateway. This is the model adopted in the
Catania Science Gateway Framework~\cite{Ardizzone2012} (see specific
documentation on
workflows\footnote{\url{http://bit.ly/1oQrzvQ}})
and LONI Pipeline Environment~\cite{dinov2009efficient}.\todo{Add better justification}

\begin{figure}
\centering
\def\svgwidth{0.5\columnwidth}
\input{images/tight.pdf_tex}
\caption{Tight integration}
\label{archi:tight}
\end{figure}

@@

See Figure~\ref{archi:service}. The workflow engine is available
externally to the science gateway. The science gateway controls it
through a specific interaction that might be implemented as a
web-service call (e.g. RESTful or SOAP), as a command-line or as any
other method that separates the workflow engine from the science
gateway. The workflow engine might be invoked either as a black box
that completely masks the infrastructure and workflow activities, or
as a grey box that allows for some interaction with them.  The
workflow engine is responsible for controlling the tasks on the
infrastructure, and performing the required data transfers to execute
them. User data is usually managed through the science gateway,
although it might as well be delivered by the workflow engine directly
to the user.

This architecture is largely adopted, in systems such as Apache
Airvata~\cite{marru2011apache}, Vine
Toolkit~\cite{DBLP:journals/scpe/SzejnfeldDKKKKLPTWDNW10}, Virtual
Imaging Platform~\cite{GLAT-13}, the WS-PGRADE/gUSE
framework~\cite{Kacsuk2012} and the numerous science gateway instances
that use it~\cite{kacsuk2014science}.
\begin{figure}
\centering
\def\svgwidth{0.5\columnwidth}
\input{images/service.pdf_tex}
\caption{Service integration}
\label{archi:service}
\end{figure}

@@

See Figure~\ref{archi:sub-task}. The science gateway is responsible to
execute the tasks on the infrastructure, dealing with heterogeneous
batch managers and meta-schedulers. The workflow engine is a
particular task that can submit sub-tasks to the science gateway. The
workflow engine keeps track of the dependencies between the sub-tasks
but their execution is delegated to the science gateway. The science
gateway may implement mechanisms to deal with task dependencies such
as basic dependency lists as available in most batch managers. It has
no global vision of the workflow. The fact that tasks are submitted
through the science gateway does not restrict the range of possible
execution models. Pilot jobs, for instance, can still be used.

This model is implemented in CBRAIN~\cite{SHER-14} where it is used to
integrate the PSOM workflow engine and the FSL toolkit through its
\texttt{fsl\_sub} tool. In particular, the CBRAIN-PSOM integration is
described in~\cite{GLAT-16} and uses an agent computing model (a.k.a pilot
jobs).
\begin{figure}
\centering
\def\svgwidth{0.5\columnwidth}
\input{images/sub-task.pdf_tex}
\caption{Sub-tasking}
\label{archi:sub-task}
\end{figure}

\subsection{Pool model}

See Figure~\ref{archi:agent}. Workflows are submitted by the science
gateway to a pool to which agents connect asynchronously to retrieve
and execute workflows (interaction \texttt{f}). Agents may be started according to various
policies, for instance to ensure load balancing. Agents may wrap
different types of workflow engines. This model was implemented in the
SHIWA pool~\cite{ROGE-13}.
\begin{figure}
\centering
\def\svgwidth{0.8\columnwidth}
\input{images/agent.pdf_tex}
\caption{Pool model}
\label{archi:agent}
\end{figure}

@@

\paragraph{Service}
+ Easy to implement in the SG (just call the service)
- Load-balancing between services is difficult
(refer to VIP's experience).
- Need to wrap workflow engines in
services.
- Need to know where to submit each workflow, e.g. when
multiple workflows are involved.
- The science gateway has no
knowledge of the detailed progress of the workflow, in particular
job-level information (statuses, stdout, stderr). If we wanted to do
that, specific feedback channels have to be developed, which requires more work (see VIP).

\paragraph{Sub-task}

\paragraph{Pool}
